Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-02-19 resex.tex
Intermediate step in revising Section II on HEL

@@

Electron lenses are based upon continuous or pulsed low energy,
magnetically confined electron
beams~\cite{Shiltsev:elens-book:2016}. The electron beam is generated
in an electron gun, guided and confined by strong solenoids and
finally dumped in a collector. As an example, the conceptual design of
the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.

\begin{figure}[b]
	\includegraphics[width=0.8\linewidth]{kick_hel_lhc_no_grid}% Here is how to import EPS art
	\caption{\label{fig:hel_field} Illustration of the hollow electron beam distribution (blue), the kick experienced by the proton beam (red) and the collimators (gray).}
The circulating beam, in case of the LHC the proton beam, is then affected by the electromagnetic field of the electron beam. For the application of active halo control, the electron beam needs to generate an electromagnetic field at the location of the halo particles while leaving the core unperturbed. This field distribution can be achieved by using a uniform hollow distribution in radius $r=\sqrt{x^2+y^2}$ with inner radius $R_1$ and outer radius $R_2$. In this case, the circulating proton beam experiences a radial kick $\theta(r)$
\begin{equation}\label{eq:field_1}
\theta(r)=\frac{f(r)}{(r/R_2)}\cdot \theta_{\rm max},
where $f(r)$ is a shape function with
\begin{equation}\label{eq:field_2}
f(r) =
\begin{cases} 0 &,\quad r< R_1,\\
\frac{r^2-R_1^2}{R_2^2-R_1^2} &,\quad R_1 \leq r < R_2,\\
1 &,\quad R_2 \leq r
\end{cases}
and $\theta_{\rm max}=\theta(R_2)$ is the maximum kick angle given by
\begin{equation}\label{hel_kick_max}
\theta_{\rm max} = \theta(R_2) = \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot \left(\frac{p_0}{q}\right)_p \cdot \beta_e\beta_p c^2}\cdot\frac{1}{R_2},
with $L$ the length of the HEL, $I_T$ the total electron beam current, $\beta_{e}$ and $\beta_{p}$ the relativistic $\beta$ of electron and proton beam, $\frac{q}{p_0}=\left(B\rho\right)_p$ the magnetic rigidity for the proton beam reference particle, $c$ the speed of light and $\epsilon_0$ the vacuum permittivity. The distribution together with the resulting kick are illustrated in Fig.~\ref{fig:hel_field}. The $\pm$-sign in Eq.~\ref{hel_kick_max} represents the two cases of the electron beam traveling in the direction of the proton beam ($v_e v_p>0$) leading to ``$-$" or in the opposite direction ($v_e v_p<0$) leading to ``$+$". For hollow electron beam collimation, the beams are chosen to counterrotate. Assuming HL-LHC and HEL design parameters (see Table~\ref{tab:hllhc_param}--\ref{tab:hel_param}) the maximum kick of the HEL is:
\begin{equation}\label{eqn:helkick}
\theta_{\rm max,B1} = 392~\rm{nrad}
%\begin{eqnarray}
%\theta_{\rm max,B1} = 392~\rm{nrad},\\
%\theta_{\rm max,B2} = 341~\rm{nrad}.
%\end{eqnarray}
for an inner radius of $R_1=4\sigma_p$, outer radius $R_2=7.5\sigma_p$, peak current of $I_e=$\SI{5.0}{A} and Beam~1 of the LHC. Similar values are obtained for Beam~2.
\begin{table}[t]
	\caption{\label{tab:hllhc_param}%
		HL-LHC design parameters at top energy \cite{hlcdr} and parameters relevant in connection with the HEL. Optics parameters at HEL are based on a position of the HEL of $-40$~m for Beam~1 (B1) and $+40$~m for Beam~2 (B2) from IP4 using HL-LHC optics V1.3 with $\beta^{*}=0.15$~m \cite{hlv13}.
	}
	\begin{ruledtabular}
		\begin{tabular}{lccc}
			Beam parameters & Value(B1) & Value(B2) & Unit\\
			\colrule
			Beam energy  $E_{p}$  &  \multicolumn{2}{c}{7} & TeV\\
			Number of bunches $n_b$ & \multicolumn{2}{c}{2748} & - \\
			Number of particles per bunch $N_b$ & \multicolumn{2}{c}{$2.2\times 10^{11}$} & -\\
			Normalized emittance $\epsilon_{N,x/y}$ & \multicolumn{2}{c}{2.5} & $\mu$m\\
			Bunch spacing & \multicolumn{2}{c}{25} & ns\\
			\colrule
			\multicolumn{4}{l}{Optics paramters at HEL (Beam~1) \footnote{As the Twiss parameters at IP4 do not change during the entire squeeze, and IP4 and the HEL are only separated by a drift space, the Twiss parameters stay constant also at the HEL during the entire squeeze.}} \\
			\colrule
			$\beta_{x}$ at HEL  & 197.5 & 280.6 & m\\
			$\beta_{y}$ at HEL & 211.9 & 262.6 m\\
			Dispersion $D_{x}$ at HEL & 0.0& 0.0 & m\\
			Dispersion $D_{y}$ at HEL & 0.0& 0.0 & m\\
			Proton beam size $\sigma_{p,x}$ at HEL & 0.26 & 0.31& mm \\			
			Proton beam size $\sigma_{p,y}$ at HEL & 0.27 & 0.30 &mm \\
			\multicolumn{4}{l}{scraping position}\\ \hspace{1cm}$\sigma_{p}=\max(\sigma_{p,x},\sigma_{p,y})$ & 0.27& 0.31 & mm\\
		\end{tabular}
	\end{ruledtabular}
\end{table}
\begin{table}[t]
	\caption{\label{tab:hel_param}%
		HL-LHC hollow electron lens parameters as in~\cite{hel_cdr}.
	}
	\begin{ruledtabular}
		\begin{tabular}{lcc}
			Geometry & Value& Unit\\
			\colrule
			Length $L$    &  3 & m\\
			% 3-8 sigmap with 3.5 um -> 3.55-9.47 for 2.5 um emittance
			Desired range of scraping positions & 3.5-9.5 &$\sigma_p$\\
			\colrule
			Magnetic fields & & \\
			\colrule
			Gun solenoid, $B_g$ & 0.2-0.4 & T\\
			Main solenoid, $B_m$ & 2-6 & T\\
			Collector solenoid, $B_c$ & 0.2-0.4 & T\\
			Compression factor ($k=\sqrt{B_m/B_g}$) & 2.2-5.5 & -\\
			\colrule
			Electron gun & & \\
			\colrule
			Peak yield $I_e$ at 10~keV & 5.0 & A\\
			Gun perveance $P$ & 5 & $\mu$perv\\
			Inner/outer cathode radius, $R_1/R_2$ & 6.75/12.7 & mm\\
			\colrule
			High voltage modulator & & \\
			\colrule
			Cathode-anode voltage & 10.0 & kV\\
			Rise time (10\%-90\%) & 200 & ns \\
			Repetition rate & 35 & kHz
		\end{tabular}
	\end{ruledtabular}
\end{table}

>>>>

\begin{figure}
  \includegraphics[width=0.8\columnwidth]{kick_hel_lhc_no_grid}
  \caption{Illustration of the hollow electron beam charge
    distribution (blue), of the magnitude of the transverse kick
    experienced by the proton beam (red), and of the position of the
    primary collimators (gray).}
  \label{fig:hel_field}
\end{figure}

\begin{table}
  \caption{HL-LHC design parameters at top energy~\cite{hlcdr} and
    parameters relevant to the HEL. Optics parameters at the HEL are
    based on a position of $-$40~m for Beam~1 (B1) and $+$40~m for
    Beam~2 (B2) from the interaction point IP4, using HL-LHC optics
    V1.3 with $\beta^{*} = \q{0.15}{m}$~\cite{hlv13}.}
  \label{tab:hllhc_param}
  \begin{ruledtabular}
    \begin{tabular}{lccc}
      Beam parameters & Value(B1) & Value(B2) & Unit\\
      \colrule
      Beam energy  $E_{p}$  &  \multicolumn{2}{c}{7} & TeV\\
      Number of bunches $n_b$ & \multicolumn{2}{c}{2748} & - \\
      Number of particles per bunch $N_b$ & \multicolumn{2}{c}{$2.2\times 10^{11}$} & -\\
      Normalized emittance $\epsilon_{N,x/y}$ & \multicolumn{2}{c}{2.5} & $\mu$m\\
      Bunch spacing & \multicolumn{2}{c}{25} & ns\\
      \colrule
      \multicolumn{4}{l}{Optics paramters at HEL (Beam~1) \footnote{As the Twiss parameters at IP4 do not change during the entire squeeze, and IP4 and the HEL are only separated by a drift space, the Twiss parameters stay constant also at the HEL during the entire squeeze.}} \\
      \colrule
      $\beta_{x}$ at HEL  & 197.5 & 280.6 & m\\
      $\beta_{y}$ at HEL & 211.9 & 262.6 m\\
      Dispersion $D_{x}$ at HEL & 0.0& 0.0 & m\\
      Dispersion $D_{y}$ at HEL & 0.0& 0.0 & m\\
      Proton beam size $\sigma_{p,x}$ at HEL & 0.26 & 0.31& mm \\		
      Proton beam size $\sigma_{p,y}$ at HEL & 0.27 & 0.30 &mm \\
      \multicolumn{4}{l}{scraping position}\\ \hspace{1cm}$\sigma_{p}=\max(\sigma_{p,x},\sigma_{p,y})$ & 0.27& 0.31 & mm\\
    \end{tabular}
  \end{ruledtabular}
\end{table}

\begin{table}
  \caption{HL-LHC hollow electron lens parameters as defined in
    Ref.~\cite{hel_cdr}.}
  \label{tab:hel_param}
  \begin{ruledtabular}
    \begin{tabular}{lcc}
      Geometry & Value& Unit\\
      \colrule
      Length $L$    &  3 & m\\
      % 3-8 sigmap with 3.5 um -> 3.55-9.47 for 2.5 um emittance
      Desired range of scraping positions & 3.5-9.5 &$\sigma_p$\\
      \colrule
      Magnetic fields & & \\
      \colrule
      Gun solenoid, $B_g$ & 0.2-0.4 & T\\
      Main solenoid, $B_m$ & 2-6 & T\\
      Collector solenoid, $B_c$ & 0.2-0.4 & T\\
      Compression factor ($k=\sqrt{B_m/B_g}$) & 2.2-5.5 & -\\
      \colrule
      Electron gun & & \\
      \colrule
      Peak yield $I_e$ at 10~keV & 5.0 & A\\
      Gun perveance $P$ & 5 & $\mu$perv\\
      Inner/outer cathode radius, $R_1/R_2$ & 6.75/12.7 & mm\\
      \colrule
      High voltage modulator & & \\
      \colrule
      Cathode-anode voltage & 10.0 & kV\\
      Rise time (10\%-90\%) & 200 & ns \\
      Repetition rate & 35 & kHz
    \end{tabular}
  \end{ruledtabular}
\end{table}

@@

Electron lenses are based upon continuous or pulsed low-energy,
magnetically confined electron beams~\cite{Shiltsev:PRSTAB:2008,
  Fischer:PRL:2015, Shiltsev:elens-book:2016}. The electron beam is
generated in an electron gun, guided and confined by strong solenoids
and finally dumped in a collector. As an example, the conceptual
design of the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.

The circulating beam (protons in the LHC case) is affected by the
electromagnetic field of the electron beam. For the application of
active halo control, the electron beam needs to generate an
electromagnetic field only at the location of the halo particles. This
field distribution can be achieved, for instance, by using a hollow
charge distribution in radius $r=\sqrt{x^2+y^2}$, uniformly
distributed between inner radius $R_1$ and outer radius $R_2$
(Fig.~\ref{hel_field}). In this case, the circulating proton beam
experiences the following radial kick $\theta(r)$:
%
\begin{equation}
  \label{eq:field_1}
  \theta(r)=\frac{f(r)}{(r/R_2)}\cdot \theta_{\rm max},
\end{equation}
%
where $f(r)$ is a shape function with
%
\begin{equation}
  \label{eq:field_2}
  f(r) =
  \begin{cases}
    0 &,\quad r< R_1,\\
    \frac{r^2-R_1^2}{R_2^2-R_1^2} &,\quad R_1 \leq r < R_2,\\
    1 &,\quad R_2 \leq r
  \end{cases}
\end{equation}
%
and $\theta_{\rm max}=\theta(R_2)$ is the maximum kick angle given by
%
\begin{equation}
  \label{hel_kick_max}
  \theta_{\rm max} = \theta(R_2) =
  \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot
    \left(p_0/q\right)_p \cdot \beta_e\beta_p
    c^2}\cdot\frac{1}{R_2},
\end{equation}
%
with $L$ the length of the HEL, $I_T$ the total electron beam current,
$\beta_{e}$ and $\beta_{p}$ the relativistic velocity parameters of
electrons and protons, $\left(p_0/q\right)_p = \left(B\rho\right)_p$
the magnetic rigidity for the proton beam reference particle, $c$ the
speed of light and $\epsilon_0$ the vacuum permittivity. The
$\pm$-sign in Eq.~\ref{hel_kick_max} represents the two cases of the
electron beam traveling in the direction of the proton beam
($v_e v_p>0$) leading to ``$-$" or in the opposite direction
($v_e v_p<0$) leading to ``$+$". For hollow electron beam collimation,
electrons and protons are chosen to counterrotate, so that the
magnetic and electric kicks add up. (For simplicity, the dependence of
the electron axial velocity on radius is neglected in
Eq.~\ref{hel_kick_max}.)

In the case of HL-LHC HEL design parameters
(Table~\ref{tab:hllhc_param} and Table~\ref{tab:hel_param}), the
maximum kick is:
%
\begin{equation}
  \label{eqn:helkick}
  \theta_{\rm max,B1} = 392~\rm{nrad}
\end{equation}
%
for an inner radius of $R_1 = 4\sigma_p$, outer radius $R_2 =
7.5\sigma_p$, peak current of $I_e = \q{5.0}{A}$, using the Beam~1
lattice of the LHC. Similar values are obtained for Beam~2.
